---
title: "ESM 204 HW4"
subtitle: "A Climate Change Model"
author: "Cameryn Brock and Allison Hacker"
date: "5/22/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```

```{r, echo = FALSE}
library(tidyverse)
```

## 1. Plots

### a) Temperature over time

```{r}
years <- data.frame(
  year = seq(from = 0, to = 200, by = 1))

# temperature function
temp_fun = function(t, T){
  temp = pmin(t*T/100,T)
  }

temp_time <- years %>% 
  mutate(none = temp_fun(t = year, T = 0),
         moderate = temp_fun(t = year, T = 2),
         extreme = temp_fun(t = year, T = 8)) %>% 
  pivot_longer(2:4,
               values_to = "temp",
               names_to = "climate_change")
```


```{r, echo = FALSE}
ggplot(temp_time, aes(x = year,
                      y = temp,
                      color = climate_change)) +
  geom_line(show.legend = FALSE) + 
  scale_color_manual(values = c("indianred", "goldenrod", "cadetblue")) +
  theme_minimal() + 
  theme(plot.margin = unit(c(1, 4, 1, 1), "lines")) +
  coord_cartesian(clip = "off") +
  geom_label(x = 201, y = 8, size = 3, 
             hjust = 0, label.size = 0,
             label = "Extreme Climate \nChange (CC)",
             color = "indianred") +
  geom_label(x = 201, y = 2, size = 3, 
             hjust = 0, label.size = 0,
             label = "Moderate CC",
             color = "goldenrod") +
  geom_label(x = 201, y = 0, size = 3, 
             hjust = 0, label.size = 0,
             label = "No CC",
             color = "cadetblue") +
  labs(x = "Year",
       y = "Temperature",
       color = "Climate Change \nScenario")
```

### b) Consumption over time

```{r}
# economic activity retained function
econ_fun <- function(temp, B = 0.05){
  econ = exp(-B*temp^2)
  }

econ_time <- temp_time %>% 
  mutate(econ = econ_fun(temp = temp))
 
# total consumption function
consum_fun <- function(econ, g = 0.01, t){
  consum = econ*exp(g*t)
  }  

consum_time <- econ_time %>% 
  mutate(consum = consum_fun(t = year, econ = econ))
```


```{r, echo = FALSE}
ggplot(consum_time, aes(x = year,
                      y = consum,
                      color = climate_change)) +
  geom_line(show.legend = FALSE) + 
  scale_color_manual(values = c("indianred", "goldenrod", "cadetblue")) +
  theme_minimal() + 
  theme(plot.margin = unit(c(1, 4, 1, 1), "lines")) +
  coord_cartesian(clip = "off") +
  geom_label(x = 201, y = 0.35, size = 3, 
             hjust = 0, label.size = 0,
             label = "Extreme CC",
             color = "indianred") +
  geom_label(x = 201, y = 6.05, size = 3, 
             hjust = 0, label.size = 0,
             label = "Moderate CC",
             color = "goldenrod") +
  geom_label(x = 201, y = 7.4, size = 3, 
             hjust = 0, label.size = 0, 
             label = "No Climate \nChange (CC)",
             color = "cadetblue") +
  labs(x = "Year",
       y = "Total Consumption",
       color = "Climate Change \nScenario")
```

### c) Undiscounted utility over time

```{r}
# society utility function
utility_fun <- function(C, n = 0.5){
  utility = ((C^(1-n))/(1-n))
  }

utility_time <- consum_time %>% 
  mutate(utility = utility_fun(C = consum))
```

```{r, echo = FALSE}
ggplot(utility_time, aes(x = year,
                      y = utility,
                      color = climate_change)) +
  geom_line(show.legend = FALSE) + 
  scale_color_manual(values = c("indianred", "goldenrod", "cadetblue")) +
  theme_minimal() + 
  theme(plot.margin = unit(c(1, 4, 1, 1), "lines")) +
  coord_cartesian(clip = "off") +
  geom_label(x = 201, y = 1.1, size = 3, 
             hjust = 0, label.size = 0,
             label = "Extreme CC",
             color = "indianred") +
  geom_label(x = 201, y = 4.9, size = 3, 
             hjust = 0, label.size = 0, 
             label = "Moderate CC", 
             color = "goldenrod") +
  geom_label(x = 201, y = 5.45, size = 3, 
             hjust = 0, label.size = 0,
             label = "No Climate \nChange (CC)", 
             color = "cadetblue") +
  scale_y_continuous(limits = c(0, 6)) +
  labs(x = "Year",
       y = "Undiscounted Utility",
       color = "Climate Change \nScenario")
```

## 2. Analysis

### a) Discounted Utility

```{r}
# discount rate function
disc_fun <- function(y = 0.005, n = 0.5, g = 0.01){
  disc = y + n*g
}

# T = 4.4
disc_4.4 <- years %>% 
  mutate(temp = temp_fun(t = year, T = 4.4),
         econ = econ_fun(temp = temp),
         consum = consum_fun(t = year, econ = econ),
         utility = utility_fun(C = consum),
         disc_utility = utility/((1 + disc_fun())^year))

# NPV under T = 4.4 
npv_4.4 <- sum(disc_4.4$disc_utility)
npv_4.4

# T = 0 
disc_0 <- years %>% 
  mutate(temp = temp_fun(t = year, T = 0),
         econ = econ_fun(temp = temp),
         consum = consum_fun(t = year, econ = econ),
         utility = utility_fun(C = consum),
         disc_utility = utility/((1 + disc_fun())^year))

# NPV under T = 0 
npv_0 <- sum(disc_0$disc_utility)
npv_0

# percent loss in PV from climate change (T = 4.4)
L <- ((npv_0 - npv_4.4)/npv_0)*100
L
```

The present value utility with climate change (T = 4.4) is **198.66**. The present value utility without climate change is **255.27**. The percent loss in present value utility from claimate change (L) is **22.18%**. 

### b) Sensitivity Analyses

#### T = 4.84

```{r}
sens_T <- years %>% 
  mutate(temp = temp_fun(t = year, T = 4.84),
         econ = econ_fun(temp = temp),
         consum = consum_fun(t = year, econ = econ),
         utility = utility_fun(C = consum),
         disc_utility = utility/((1 + disc_fun())^year))

npv_sens_T <- sum(sens_T$disc_utility)

# L with T = 4.84
L_sens_T <- ((npv_0 - npv_sens_T)/npv_0)*100
L_sens_T

# change in L
L_sens_T - L
```

Increasing T by 10% results in an **increase** in the percent loss in present value utility (L) by **3.62%**.

#### g = 0.011

```{r}
sens_g <- years %>% 
  mutate(temp = temp_fun(t = year, T = 4.4),
         econ = econ_fun(temp = temp),
         consum = consum_fun(t = year, econ = econ, g = 0.011),
         utility = utility_fun(C = consum),
         disc_utility = utility/((1 + disc_fun(g = 0.011))^year))

npv_sens_g <- sum(sens_g$disc_utility)

# L with g = 0.011
L_sens_g <- ((npv_0 - npv_sens_g)/npv_0)*100
L_sens_g

# change in L 
L_sens_g - L
```
Increasing g by 10% results in a **decrease** in the percent loss in present value utility (L) by **0.03%**.

#### n = 0.55

```{r}
sens_n <- years %>% 
  mutate(temp = temp_fun(t = year, T = 4.4),
         econ = econ_fun(temp = temp),
         consum = consum_fun(t = year, econ = econ),
         utility = utility_fun(C = consum, n = 0.55),
         disc_utility = utility/((1 + disc_fun(n = 0.55))^year))

npv_sens_n <- sum(sens_n$disc_utility)

# L with n = 0.55
L_sens_n <- ((npv_0 - npv_sens_n)/npv_0)*100
L_sens_n

# change in L 
L_sens_n - L
```

Increasing n by 10% results in a **decrease** in the percent loss in present value utility (L) by **4.51%**.

#### B = 0.055

```{r}
sens_b <- years %>% 
  mutate(temp = temp_fun(t = year, T = 4.4),
         econ = econ_fun(temp = temp, B = 0.055),
         consum = consum_fun(t = year, econ = econ),
         utility = utility_fun(C = consum),
         disc_utility = utility/((1 + disc_fun())^year))

npv_sens_b <- sum(sens_b$disc_utility)

# L with B = 0.055
L_sens_b <- ((npv_0 - npv_sens_b)/npv_0)*100
L_sens_b

# change in L 
L_sens_b - L
```

Increasing B by 10% results in an **increase** in the percent loss in present value utility (L) by **1.76%**.

### c)

$$U(C_{4.4})=\frac{(\theta*C_0)^{1-n}}{1-n}$$


$$U(C_{4.4})=\theta^{1-n}*\frac{C_0^{1-n}}{1-n}$$
$$U(C_{4.4})=\theta^{1-n}*U(C_0)$$
$$\theta^{1-n} = \frac{U(C_{4.4})}{U(C_0)}$$
Assuming n = 0.5,

$$\theta = (\frac{U(C_{4.4})}{U(C_0)})^2$$


```{r}
# Without discounting

U_4.4 <- sum(disc_4.4$utility)
U_0 <- sum(disc_0$utility)

theta = (U_4.4/(U_0))^2
theta
```

```{r}
# With discounting

U_4.4_disc <- sum(disc_4.4$disc_utility)
U_0_disc <- sum(disc_0$disc_utility)

theta_disc = (U_4.4_disc/(U_0_disc))^2
theta_disc
```


### d)

```{r}



```

