---
title: "ESM204 Homework 4"
author: "Allison Hacker and Cameryn Brock"
date: "5/20/2020"
output: html_document
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, include=FALSE}
library(tidyverse)
```


### Equations and Plots
```{r}
# temperature

#define variables
sims = 201
temps <- rep(NA, sims)
temps_0 <- rep(NA, sims)
temps_2 <- rep(NA, sims)
temps_8 <- rep(NA, sims)
t <- seq(from = 0, to = 200, by = 1)

# temperature function
temp_fun<- function(T, x){
  temp = min((T*x)/100, T)
}

# create temp vectors for CC scenarios
for(i in 1:sims){
  temps_0[i] <- temp_fun(0, i-1)
}

for(i in 1:sims){
  temps_2[i] <- temp_fun(2, i-1)
}

for(i in 1:sims){
  temps_8[i] <- temp_fun(8, i-1)
}

# combine into dataframe
temps_data_raw <- data.frame(t, temps_0, temps_2, temps_8)
temps_data <- temps_data_raw %>% 
  pivot_longer(cols = starts_with("temps"), names_to = "cc_scenario", values_to = "temp")

# plot temps
ggplot(data = temps_data, aes(x = t, 
                              y = temp, 
                              color = cc_scenario))+
  geom_line()

```


```{r}
# economic activity
K <- rep(NA, sims)
K_0<- rep(NA, sims)
K_2<- rep(NA, sims)
K_8<- rep(NA, sims)

B = 0.05

econ_fun <- function(temp, B = 0.05){
  K = exp(-B*temp^2)
}

for(i in 1:sims){
   K_0[i] <- econ_fun(temps_0[i])
}

for(i in 1:sims){
   K_2[i] <- econ_fun(temps_2[i])
}

for(i in 1:sims){
   K_8[i] <- econ_fun(temps_8[i])
}

```

```{r}
# consumption
C <- rep(NA, sims)
C_0<- rep(NA, sims)
C_2<- rep(NA, sims)
C_8<- rep(NA, sims)

con_fun <- function(K, x, g = 0.01){
  C = K*exp(g*x)
}

 for(i in 0:sims){
   C_0[i] <- con_fun(K_0[i], i-1)
 }

for(i in 0:sims){
   C_2[i] <- con_fun(K_2[i], i-1)
}

for(i in 0:sims){
   C_8[i] <- con_fun(K_8[i], i-1)
}

# combine into dataframe
con_data_raw <- data.frame(t, C_0, C_2, C_8)
con_data <- con_data_raw %>% 
  pivot_longer(cols = starts_with("C"), names_to = "cc_scenario", values_to = "C")

# plot K
ggplot(data = con_data, aes(x = t, y = C, color = cc_scenario))+
  geom_line()
```

```{r}
# Utility
U <- rep(NA, sims)
U_0<- rep(NA, sims)
U_2<- rep(NA, sims)
U_8<- rep(NA, sims)

u_fun <- function(C, n = 0.5){
  U = (C^(1-n))/(1-n)
}

for(i in 0:sims){
   U_0[i] <- u_fun(C_0[i])
}

for(i in 0:sims){
   U_2[i] <- u_fun(C_2[i])
}

for(i in 0:sims){
   U_8[i] <- u_fun(C_8[i])
}

```

```{r}
# Discounting
d = 0.005
n = 0.5
g = 0.01

r_fun <- function(d = 0.005, n = 0.5, g = 0.01){
  r = d + n*g
}

discount <- function(U, r){
  pv <- rep(NA, sims)
  for(i in 1:sims){
    pv[i] = U[i]/(1+r)^(i-1)
  }
  pv_total = sum(pv)
  return(pv_total)
}
```

### Analysis

##### Part A
```{r}
# Function for percent change
p_change <- function(L0, L_new){
  pc <- (L_new-L0)/L0
  return(pc*100)
}

# Function for loss in present value of utility

loss_fun <- function(T, B = 0.05, d=0.005, n = 0.5, g=0.01){
# Climate Change
for(i in 1:sims){
  temps[i] <- temp_fun(T, i-1)}

for(i in 1:sims){
  K[i] <- econ_fun(temps[i], B = B)}

for(i in 1:sims){
  C[i] <- con_fun(K[i], i-1, g = g)}

for(i in 1:sims){
  U[i] <- u_fun(C[i], n = n)}

r <- r_fun(d = d, n = n, g = g)

cc_utility <- discount(U, r)

# No climate change
T <- 0

for(i in 1:sims){
  temps[i] <- temp_fun(T, i-1)}

for(i in 1:sims){
  K[i] <- econ_fun(temps[i])}

for(i in 1:sims){
  C[i] <- con_fun(K[i], i-1, g = g)}

for(i in 1:sims){
  U[i] <- u_fun(C[i], n = n)}

utility <- discount(U, r)

L <- abs(p_change(L0 = utility, L_new = cc_utility))
return(L)
}

L_test <- loss_fun(T = 4.4)
```


##### Part B
```{r}

T <- 4.4
T10 <- T*1.1
B10 <- B*1.1
n10 <- n*1.1
g10 <- g*1.1

L_T <- loss_fun(T = T10)
L_B <- loss_fun(T = 4.4, B = B10)
L_n <- loss_fun(T = 4.4, n = n10)
L_g <- loss_fun(T = 4.4, g = g10)

pcT <- p_change(L0 = L_test, L_new = L_T)
pcB <- p_change(L0 = L_test, L_new = L_B)
pcn <- p_change(L0 = L_test, L_new = L_n)
pcg <- p_change(L0 = L_test, L_new = L_g)

```


##### Part C

Back to the original parameters, suppose we could completely prevent climate
change from occurring (so T = 0 instead of T = 4.4) but doing so would require
giving up a fraction theta of consumption every year for the next 200 years. What
is the maximum value of theta society would be willing to endure every year to
completely prevent climate change? Call this theta*.

```{r}
# what fraction of consumption would society be willing to "pay" every year for 200 years to completely prevent climate change?

# sum(utility (T = 0, C*theta)) <= sum(utility (T= 4.4))

temps_cc <- rep(NA, sims)
K_cc <- rep(NA, sims)
C_cc <- rep(NA, sims)
U_cc <- rep(NA, sims)

T = 4.4

for(i in 1:sims){
  temps_cc[i] <- temp_fun(T, i-1)}

for(i in 1:sims){
  K_cc[i] <- econ_fun(temps_cc[i], B = B)}

for(i in 1:sims){
  C_cc[i] <- con_fun(K_cc[i], i-1)}

for(i in 1:sims){
  U_cc[i] <- u_fun(C_cc[i], n = n)}

r <- r_fun(d = d, n = n, g = g)

cc_utility <- sum(U_cc)

# Find theta

T <- 0

for(i in 1:sims){
  temps[i] <- temp_fun(T, i-1)}

for(i in 1:sims){
  K[i] <- econ_fun(temps[i])}

for(i in 1:sims){
  C[i] <- con_fun(K[i], i-1, g = g)}

u_theta <- rep(NA, sims)
u_fun_theta <- function(C, n = 0.5){
  u_theta = C^(1-n)
}

for(i in 1:sims){
  u_theta[i] <- u_fun_theta(C_0[i])}

# U_cc = theta^0.5 * sum(u_theta)*2
# theta^0.5 = U_cc/(2*sum(u_theta))
# theta = (U_cc/(2*sum(u_theta)))^2

theta = (cc_utility/(2*sum(u_theta)))^2

# find utility with no climate change and theta
C_theta <- C_0*theta
u_theta_test<- rep(NA, sims)

for(i in 1:sims){
  u_theta_test[i] <- u_fun(C_theta[i])}
sum(u_theta_test)

#cc_utility_disc <- discount(U_cc, r)
#theta_utility_disc <- discount(u_theta_test, r)

```

theta* = 0.493


##### Part D

Suppose we are uncertain about T, but it has the following probability distribution:
T = 2 (with probability .2), T = 4 (with probability .5), and T = 6 (with
probability .3). Calculate theta under uncertainty over T.

```{r}

temps_4 <- rep(NA, sims)
temps_6 <- rep(NA, sims)

# create temp vectors for CC scenarios
for(i in 1:sims){
  temps_4[i] <- temp_fun(4, i-1)}

for(i in 1:sims){
  temps_6[i] <- temp_fun(6, i-1)}

K_4<- rep(NA, sims)
K_6<- rep(NA, sims)

for(i in 1:sims){
   K_4[i] <- econ_fun(temps_4[i])}

for(i in 1:sims){
   K_6[i] <- econ_fun(temps_6[i])}

C_4<- rep(NA, sims)
C_6<- rep(NA, sims)

 for(i in 0:sims){
   C_4[i] <- con_fun(K_4[i], i-1)}

for(i in 0:sims){
   C_6[i] <- con_fun(K_6[i], i-1)}

U_4<- rep(NA, sims)
U_6<- rep(NA, sims)

for(i in 0:sims){
   U_4[i] <- u_fun(C_4[i])}

for(i in 0:sims){
   U_6[i] <- u_fun(C_6[i])}

expected_U <- sum(U_2)*0.2 + sum(U_4)*0.5 + sum(U_6)*0.3
theta_expected = (expected_U/(2*sum(u_theta)))^2

```

theta* = 0.515